### 2013-09-05 structured_prediction/paper.tex
structured prediction: More elaborate description of learning algorithms

There are several popular algorithms to solve \Eqref{learning_equation}. We
will briefly review three standard algorithms: the one-slack and n-slack
cutting plane algorithms, a stochastic primal subgradient algorithm, and
recently proposed stochastic dual coordinate descent method.

\subsection{Stochastic Subgradient Descent}\seclabel{subgradient}
Arguably the most straight-foward way to approach \Eqref{learning_equation} is
using subgradient descent.  In light of the complexity of solving the
loss-augmented prediction problem in \Eqref{loss_augmentation}, is is natural
to work in a stochastic setting (see \citet{ratliff2007online}).
Given a model through $\phi$ and a set of parameters $\theta$, a subgradient
considering a single training example $(x\hoch{i}, y\hoch{i})$
can be computed simply by solving the loss-augmented prediction problem:
\begin{align}
    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\\
    &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y)
\end{align}
The most commonly used update has the simple form
\begin{equation}
    \theta \leftarrow (1 - \eta_t) \theta - \eta_t C \left [\phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i})\right]
\end{equation}
Where $\eta_t$ is a sequence of learning rates.
%Theoretical convergence guarantees of \ldots can be given whenever \ldots,
but in practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
Many practitioners adopted the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{pegasos}:
%\begin{equation}

%\end{equation}

%\citep{ratliff2007online} derive convergence rates for a stochastic version of the subgradient of  Equation~\ref{mainequation}:
%\begin{equation}\label{subgradient}
    %g := \phi(x_i, y) - \phi(x_i, y_i) + \theta / C
%\end{equation}
%leading to a simple iterative optimization.

@@

The n-slack cutting plane method~\citep{tsochantaridis2006large} reformulates \Eqref{learning_equation}
into a quadratic objective with a combinatorial number of constraints:
\begin{align}\label{eq:oneslack}
    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \sum_{i=1}^k \xi_i\\
    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \mathcal{Y}^k:\\
    &\theta^T \sum_{i=1}^n [\phi(x\hoch{i}, y\hoch{i}) - \phi(x\hoch{i},
        \hat{y}\hoch{i})] \geq \sum_{i=1}^k \Delta(y\hoch{i}, \hat{y}\hoch{i})
            - \xi
\end{align}
As is not feasible to deal with all constraints, only a working set $\W$ of active constraints
is maintained, using the cutting plane method. The algorithm starts with an empty working set,
and in each iteration adds the most violated constraint to $\W$. Then, the quadratic program is solved
again, with the new set of constraints.
The algorithm terminates when no constraint can be found that is more violated then $\epsilon$,
which guarantees a suboptimality of at most $\epsilon$.

The complete procedure is described in Algorithm~\ref{alg_n_slack}.

@@

The one-slack cutting plane method~\citep{joachims2009cutting} solves the
following reformulation of Equation~\Eqref{learning_equation}:
\begin{align}\label{eq:oneslack}
    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \xi\\
    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
        &\theta^T \sum_{i=1}^n [\phi(x^i, y^i) - \phi(x^i,
            \hat{y}^i)] \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
            - \xi
\end{align}
Informally, the one-slack formulation corresponds to joining all training
samples into a single training example $(\mathbf{x}, \mathbf{y})$ that has no
interactions between variables corresponding to different data points, and
then applying the n-slack algorithm.
A detailed description can be found in Algorithm~\ref{alg_one_slack}.

By construction only a single constraint is added in each iteration of
Algorithm~\ref{alg_one_slack}, leading to very small working sets $\W$.
This has the advantage of making the solution of the QP faster, as it contains
less variables. The down-side of this is that the loss-augmented prediction
problem has to be solve much more often until convergence.
%TODO convergence rate

@@

\subsection{The STLJSP algorithm}\seclabel{dual_coordinate_descent}
Very recently, two groups independently derived a very performant new algorithm.
\citet{shalev2012proximal} investigated stochastic dual coordinate descent applied
to structured support vector machines. \citet{lacoste2012block} started from the Frank-Wolfe algorithm %TODO details
and derived a block-coordinate version, where each block corresponds to the constraints associated with a single training example.
Both groups showed that closed-form line search is possible, and, surprisingly ended up with
the exact same algorithm and closely related convergence guarantees.
The algorithm processes a single training example at a time, and can be applied while
remaining completely in the primal domain, making it applicable to very large datasets.
We will refer to the algorithm as STDJSP, after the authors of the two papers, to honor
the independent development of the two views of the method.
The STDJSP algorithm has properties that are very simliar to stochastic subgradient descent,
but with two main advantages:
\begin{description}
\item[Stopping Criterion] Both views of STLJSP give rise to a dual objective, which can be used as a theoretically sound stopping criterion.
\item[Learning Rate] TODO %TODO
\end{description}

>>>>

There are several popular algorithms to solve \Eqref{learning_equation}. We
will briefly review three standard algorithms: the one-slack and n-slack
cutting plane algorithms, a stochastic primal subgradient algorithm, and
recently proposed stochastic dual coordinate descent method, which we
will now describe in detail.

% discuss online vs batch?

We will also discuss practical implications and implementation.
One particularly intereresting aspect is parallelization.
In image segmentation tasks, inference is often costly, making loss-augmented
prediction the most expensive step in learning. In \emph{batch} algorithms,
which process all training examples at once, loss-augmented inference is
embarassingly parallel, leading to a linear speed-up in terms of processes.

\subsection{Stochastic Subgradient Descent}\seclabel{subgradient}

Arguably the most straight-foward way to approach \Eqref{learning_equation} is
using subgradient descent.  In light of the complexity of solving the
loss-augmented prediction problem in \Eqref{loss_augmentation}, is is natural
to work in a stochastic setting (see \citet{ratliff2007online}).
Given a model through $\phi$ and a set of parameters $\theta$, a subgradient
considering a single training example $(x\hoch{i}, y\hoch{i})$
can be computed simply by solving the loss-augmented prediction problem:
\begin{align}
    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\\
    &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y)
\end{align}
The most commonly used update has the simple form
\begin{equation}
    \theta \leftarrow (1 - \eta_t) \theta - \eta_t C \left [\phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i})\right]
\end{equation}
Where $\eta_t$ is a sequence of learning rates.
%Theoretical convergence guarantees of \ldots can be given whenever \ldots,
but in practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
Many practitioners adopted the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{pegasos}:
\begin{equation}
    \eta_t = \frac{C}{t},
\end{equation}
which has been found to work well in many settings.
Implementation of the algorithm is straight-forward, but unfortunately
detecting convergence is often tricky.
It is possible to use mini-batches instead of processing one sample at a time
to make use of multiple processors for inference. Unfortunately this negatively
affects the number of iterations needed, and did not provide a benefit in practice.

@@

The n-slack cutting plane method~\citep{tsochantaridis2006large} reformulates \Eqref{learning_equation}
into a quadratic objective with a combinatorial number of constraints:
\begin{align}\label{eq:oneslack}
    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \sum_{i=1}^k \xi_i\\
    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \mathcal{Y}^k:\\
    &\theta^T \sum_{i=1}^n [\phi(x\hoch{i}, y\hoch{i}) - \phi(x\hoch{i},
        \hat{y}\hoch{i})] \geq \sum_{i=1}^k \Delta(y\hoch{i}, \hat{y}\hoch{i})
            - \xi
\end{align}
As is not feasible to deal with all constraints, only a working set $\W$ of active constraints
is maintained, using the cutting plane method. The algorithm starts with an empty working set,
and in each iteration adds the most violated constraint to $\W$. Then, the quadratic program is solved
again, with the new set of constraints.
The algorithm terminates when no constraint can be found that is more violated then $\epsilon$,
which guarantees a suboptimality of at most $\epsilon$.
%
The complete procedure is described in Algorithm~\ref{alg_n_slack}.
The n-slack cutting plane can be understood is a sequential algorithm that processes each sample individually.
While this allows fast process of the optimization with respect to the number of calls to loss-augmented prediction,
individual iterations become more and more costly. The number of the constraints is usually a multiple of
the number of training  samples, which leads to very large QP problems, even for medium sized datasets.
This makes the algorithm often slow in practice.
%
Solving the quadratic programm can be accelerated using several techniques. We found that
agressively removing constraints that are inactive or contribute little to the solution
often makes the difference between the algorithm being practical or not.
Another possible technique is to update the quadratic program only every $r$ samples, for some
small integer $r$ (\citet{joachims2009cutting} suggest 100). While this strategy on its own
did not provide a large benefits in our experiments, it allows for parallel loss-augmented
inference on these mini-batches of size $r$.

@@

The one-slack cutting plane method~\citep{joachims2009cutting} solves the
following reformulation of Equation~\Eqref{learning_equation}:
\begin{align}\label{eq:oneslack}
    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \xi\\
    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
        &\theta^T \sum_{i=1}^n [\phi(x^i, y^i) - \phi(x^i,
            \hat{y}^i)] \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
            - \xi
\end{align}
Informally, the one-slack formulation corresponds to joining all training
samples into a single training example $(\mathbf{x}, \mathbf{y})$ that has no
interactions between variables corresponding to different data points, and
then applying the n-slack algorithm.
A detailed description can be found in Algorithm~\ref{alg_one_slack}.
%
By construction only a single constraint is added in each iteration of
Algorithm~\ref{alg_one_slack}, leading to very small working sets $\W$.
This has the advantage of making the solution of the QP faster, as it contains far
less variables than in the $n$-slack algorithm. The down-side of this is that
the loss-augmented prediction problem has to be solve much more often until
convergence.
%TODO convergence rate

@@

\subsection{The STLJSP algorithm}\seclabel{dual_coordinate_descent}

Very recently, two groups independently derived a very performant new algorithm.
\citet{shalev2012proximal} investigated stochastic dual coordinate descent applied
to structured support vector machines. \citet{lacoste2012block} started from the Frank-Wolfe algorithm %TODO details
and derived a block-coordinate version, where each block corresponds to the
constraints associated with a single training example.  Both groups showed that
closed-form line search is possible, and, surprisingly ended up with the exact
same algorithm and closely related convergence guarantees.  The algorithm
processes a single training example at a time, and can be applied while
remaining completely in the primal domain, making it applicable to very large
datasets.  We will refer to the algorithm as STLJSP, after the authors of the
two papers, to honor the independent development of the two views of the
method.
%
The STDJSP algorithm has properties that are very simliar to
stochastic subgradient descent, but with two main advantages:
\begin{description}
\item[Stopping Criterion] Both views of STLJSP give rise to a dual objective,
    which can be used as a theoretically sound stopping criterion.
\item[Learning Rate] While there are several theoretical results on choosing
    the learning rate in stochastic subgradient descent, choosing a concrete
    schedule is often problematic in practice. By using analytic line-search,
    STLJSP removes the need for any learning rate parameter, making application
    of the algorithm in practice much simpler.
\end{description}

The STLJSP algorithm is usually used with weighted averaging, as this is known
to improve performance in similar stochastic algorithms, and empirically
improves convergence. % TODO weighted averaging here? or at SGD?
